Delayed effects of [responses to?] climate lead to demographic divergence in a fragmented Amazonian landscape

### 2021-05-04 doc/paper.Rmd

Third, it may be that the delayed demographic effects we observed are indirectly mediated by the effect of SPEI on other species rather than the direct effects on individual physiology [@evers2021]. For example, while droughts can kill some tropical trees immediately [@laurance2001], others take three years or more [@crileyYearProductionHigh1994]. This will potentially lead to a boom in growth and flowering years after the drought, when the canopy gaps resulting from delayed leaf drop [@janssenDroughtEffectsLeaf2021] and treefalls allow for light penetration to the forest understory [@canham1990]. Similar delayed changes in the local environment could also influence the foraging behavior of *H. acuminata*'s pollinators [@bruna2004; @stouffer1996] and seed dispersers [@uriarte2011]. While more work is needed to explain why the delayed effects of SPEI on survival and growth are greater in forest fragments than interiors, [**question about wording** - does this mean the delay is relatively greater, or the magnitude of the effect is greater?**]** it may be due to the proportionately greater amount of litterfall on edges [@vasconcelos2004] altering the abundance of pathogens or mycorrhizae.

>>>>

Third, it may be that the delayed demographic effects we observed are indirectly mediated by the effect of SPEI on other species rather than the direct effects on individual physiology [@evers2021]. For example, while droughts can kill some tropical trees immediately [@laurance2001], others take three years or more [@crileyYearProductionHigh1994]. This will potentially lead to a boom in growth and flowering years after the drought, when the canopy gaps resulting from delayed leaf drop [@janssenDroughtEffectsLeaf2021] and treefalls allow for light penetration to the forest understory [@canham1990]. Similar delayed changes in the local environment could also influence the foraging behavior of *H. acuminata*'s pollinators [@bruna2004; @stouffer1996] and seed dispersers [@uriarte2011]. While more work is needed to explain why the delayed effects of SPEI on survival and growth are greater in forest fragments than interiors, [**question about wording** - does this mean the delay is relatively greater, or the magnitude of the effect is greater?\*\*]\*\* it may be due to the proportionately greater amount of litterfall on edges [@vasconcelos2004] altering the abundance of pathogens or mycorrhizae.
